---
output:
  md_document:
    variant: gfm
html_preview: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
``` 

```{r echo=FALSE, results="hide", message=FALSE}
library("badger")
```

```{r, echo = FALSE, results='asis'}
cat(
  badge_github_actions("helseprofil/KHompare"),
  badge_codecov("helseprofil/KHompare"),
  badge_lifecycle("experimental", "blue"),
  badge_devel("helseprofil/KHompare", "blue")
)
```


# KHompare

Compare KHelse data to check for any abnormal changes.

# Installation

To install the package run

```r
if(!requirenamespace("remotes")) install.packages("remotes")
remotes::install_github("helseprofil/orgdata")
```

# Usage 

To check a cube file for *ALKOHOL*

```r
library("KHompare")
dt <- check_cube("ALKOHOL")
```

Comparison will be made on the measurements created by `LagKUBE()` function in
**KHfunction** such as:

- MEIS
- RATE
- TELLER
- SMR
- etc ..

New columns of these measurement variables will be created when running the
function `check_cube()` and they are:

- *xxx*_NUM
- *xxx*_PCT

The *xxx* is the name of the measurement variables that are checked for the
difference in change numerically and percent. They are denoted by *NUM* and
*PCT* respectively.
